Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy

### 2015-01-21 AO_bondedSi.tex
add VS20-21 IR image, add text about correlated errors,

\section{Measurement strategy, conditions, and details}
We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.  Table \ref{tab:Cary500pars} lists the typical measurement settings and tool performance.  The tool has a double monochromator, which reduces scattered light.  The vendor reported linearity exceeds 40 dB.

The Cary 5000 has four operation modes: single front, single back, double, and double reverse.  We experimented with single front and double modes.  Double mode takes a reference spectrum simultaneously with the target spectrum.  In both single and double mode, a baseline scan is taken with no sample in the measurement holder.  The baseline spectrum is the raw spectral response of the system.  We automatically divide sample spectra by the baseline spectrum.  In single beam mode, the baseline measurements must be repeated about every half hour for post-process division.  In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.  We estimated the uncertainty and measurement repeatability of single and double modes in the following way: we computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.  We found that the mean value in single mode drifted by about 0.2\% over several measurements.  The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption.  In double beam mode, the spectrograph drift was much lower.  The simultaneous reference in double beam mode divides out the 0.2\% drift in the mean transmission level attributable to lamp drift that we saw in single beam measurements.  In double beam mode, the drift in mean transmission level is 0.02\%.  These values are accurate for measurements over a half hour timescale or less.  With accuracies and drift rates at this level, the curves in Figure \ref{figAbsoluteTrans} indicate that the axial extent of gaps as small as a few nm should be measurable.

@@

Before testing the bonded pieces from Table 1 using spectrophotometry, we looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.  The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.  We used a 50 mm F/2 lens, with a fiber optic white light illumination source.  The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately 1.15$-$1.6 $\mu$m.  The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.  No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.  It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the sample when we dither its position.  We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.  The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.  As expected, the unbonded DSP wafers show no such IR bubbles.  No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.  The morphology was clear regardless of the detector response.

@@

\subsection{Spectra of direct bonded Si with known gap sizes}
Figures \ref{VS21specm1} to \ref{VS21specf1} show examples of the predicted and measured IR spectra.  Figure \ref{VS21specm1} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.  Meanwhile, Figure \ref{VS21specf1} shows a spectrum in which the measurement marginally exceeds the prediction (by about $\sim 2 \sigma$).  Other parts show less consistency with the prediction.

>>>>

\section{Measurement technique and correlated measurement errors}
We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.  Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.  The tool has a double monochromator, which reduces scattered light.  The vendor reported linearity exceeds 40 dB.

The Cary 5000 has four operation modes: single front, single back, double, and double reverse.  We experimented with single front and double modes.  Double mode takes a reference spectrum simultaneously with the target spectrum.  In both single and double mode, a baseline scan is taken with no sample in the measurement holder.  The baseline spectrum is the raw spectral response of the system.  We automatically divide sample spectra by the baseline spectrum.  In single beam mode, the baseline measurements must be repeated about every half hour for post-process division.  In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.

We estimated the uncertainty and measurement repeatability of single and double modes in the following way: we computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.  We found that the mean value in single mode drifted by about 0.2\% over several measurements.  The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption.  In double beam mode, the spectrograph drift was much lower.  The simultaneous reference in double beam mode divides out the 0.2\% drift in the mean transmission level attributable to lamp drift that we saw in single beam measurements.  In double beam mode, the drift in mean transmission level is 0.02\%.  These values are accurate for measurements over a half hour timescale or less.  Table \ref{tabCary5000pars} summarizes the amplitudes of the drifts and individual measurement errors in both single and double beam modes.\emph{Correlated errors dominate the measurement uncertainty of our transmission spectra.}  The measurement uncertainty correlations are apparent in in Figure \ref{figSiAbsorbfig}.  In some parts of the spectrum, the mean drift is about ten times larger than the standard deviation of the spectrum.  Double beam mode has about ten times lower lamp drift but suffers from other instrumental artifacts, like disjointed spectra at the wavelengths of known grating changeover points.

@@

We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.  The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.  We used a 50 mm F/2 lens, with a fiber optic white light illumination source.  The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately 1.15$-$1.6 $\mu$m.  The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.  No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.  It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the sample when we dither its position.  We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.  The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.  As expected, the unbonded DSP wafers show no such IR bubbles.  No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.  The morphology was clear regardless of the detector response.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.3\columnwidth]{figs/VS2021_IR_image.bmp}}
\caption{
\label{figVS2021_IR_image}
IR image of part VS20-21. Three rectangular mesh areas are apparent (Table \ref{tabPatternFills}).  The coarsest mesh, C, shows substructure, while the medium and fine meshes are indistinguishable at the low resolution of the image.  The three dark spots are gaps at the inferface of the direct bonded Si samples.}
\end{figure}

@@

\subsectio{Inferring axial extent of gaps from noisy transmission spectra}
We infer the axial extent of the gaps with correlated measurement uncertainties using Gaussian Processes.

\subsection{Spectra of direct bonded Si with known gap sizes}
Figures \ref{figVG03full} to \ref{figVG12_f50} show examples of the predicted and measured IR spectra.  Figure \ref{figVG03_f100} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.  Meanwhile, Figure \ref{figVG12_f50} shows a spectrum in which the measurement marginally exceeds the prediction (by about $\sim 3 \sigma$).  Other parts show less consistency with the prediction due to the difficulty of direct-bonding Si in the absence of thermal annealing.
